---
title: "Heart Disease PCA and SVM"
author: "Lucas Mendicino"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Support Vector Machine


# Load Packages
```{r}
library(tidyverse)
library(caret)
library(e1071)
library(factoextra)
library(FactoMineR)
library(PCAmixdata)
```

# Prepare the Data

Let's load the data
```{r}
# Read datasets Cleveland_hd.csv into hd_data
hd <- read.csv("~/Desktop/UCI-Heart-Disease-Prediction/heart.csv")
```

Let's look at the structure and a few rows
```{r}
str(hd)
head(hd)
```
                                            
The variables include:

- age: age in years
- sex: (1 = male; 0 = female)
- cp: chest pain type (typical angina, atypical angina, non-anginal pain, or asymptomatic angina)
- trestbps: resting blood pressure (in mm Hg on admission to the hospital)
- chol: serum cholestoral in mg/dl
- fbs: Fasting blood sugar (< 120 mg/dl or > 120 mg/dl) (1 = true; 0 = false)
- restecg: resting electrocardiographic results (normal, ST-T wave abnormality, or left ventricular hypertrophy)
- thalach: Max. heart rate achieved during thalium stress test
- exang: Exercise induced angina (1 = yes; 0 = no)
- oldpeak: ST depression induced by exercise relative to rest
- slope: Slope of peak exercise ST segment (0 = upsloping, 1 = flat, or 2 = downsloping)
- ca: number of major vessels (0-3) colored by flourosopy 4 = NA
- thal: Thalium stress test result 3 = normal; 6 = fixed defect; 7 = reversable defect 0 = NA
- target: Heart disease status 1 or 0 (0 = heart disease 1 = asymptomatic)

It is important to note that for the target variable, 0 is assigned to heart disease and 1 is asymptomatic. Let's recode this.


```{r}
hd_new <- hd %>% mutate(target = recode(target, "0" = "1", "1" = "0"))
```


Let's tidy the dataset a bit more.

```{r}
hd_model <- hd_new %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
  ) %>%
# Recode the categorical variables as factors using the dplyr library.
  mutate(
    sex = case_when(
      sex == 0 ~ "female",
      sex == 1 ~ "male"
           ),
    fbs = case_when(
      fbs == 0 ~ "<=120",
      fbs == 1 ~ ">120"
            ),
    exang = case_when(
      exang == 0 ~ "no",
      exang == 1 ~ "yes"
            ),
    cp = case_when(
      cp == 3 ~ "typical angina",
      cp == 1 ~ "atypical angina",
      cp == 2 ~ "non-anginal pain",
      cp == 0 ~ "asymptomatic angina"
          ),
    restecg = case_when(
      restecg == 0 ~ "hypertrophy",
      restecg == 1 ~ "normal",
      restecg == 2 ~ "wave abnormality"
              ),
    target = case_when(
      target == 0 ~ "asymptomatic",
      target == 1 ~ "heart-disease"
              ),
    slope = case_when(
      slope == 2 ~ "upsloping",
      slope == 1 ~ "flat",
      slope == 0 ~ "downsloping"
    ),
    thal = case_when(
      thal == 1 ~ "fixed defect",
      thal == 2 ~ "normal",
      thal == 3 ~ "reversable defect"
    ),
    sex = as.factor(sex),
    fbs = as.factor(fbs),
    exang = as.factor(exang),
    cp = as.factor(cp),
    slope = as.factor(slope),
    ca = as.factor(ca),
    thal = as.factor(thal),
    restecg = as.factor(restecg),
    target = as.factor(target)
  )

```



## PCA



```{r}
hd_model1 <- hd_model %>% mutate(cp = as.factor(cp), restecg = as.factor(restecg),
                                            exang = as.factor(exang), slope = as.factor(slope),
                                            ca = as.factor(ca), thal = as.factor(thal), fbs = as.factor(fbs),
                                            sex = as.factor(sex)) 

hd_model2 <- hd_model1 %>% select(-c(target))

str(hd_model2)
```

Using PCA of Mixed Data
```{r}
X.quanti <- splitmix(hd_model2)$X.quanti
X.quali <- splitmix(hd_model2)$X.quali

pca <- PCAmix(X.quanti, X.quali, ndim = 6, rename.level = TRUE)
summary(pca)
```

```{r}
pca$eig
```



```{r}
barplot(pca$eig[,2], ylim = c(0,20), las = 2)
```


```{r}
pca1 = data.frame(pca$ind$coord, hd_model1$target)
ggplot(pca1, aes(x = dim.1, y = dim.2, color = hd_model1.target)) + 
  geom_point() + 
  theme_bw() +
  guides(color = guide_legend(title = "Has heart \n disease?")) + 
  ggtitle("PCA of heart disease") +
  xlab(paste("Dimension 1", paste0("(", 
                                   round(pca$eig[1, 2], 2), 
                                   "%", ")"))) + 
  ylab(paste("Dimension 2", paste0("(", 
                                   round(pca$eig[2, 2], 2), 
                                   "%", ")")))
```
```{r}
pca2 = data.frame(pca$sqload)
# Dimension 1
ggplot(pca2, aes(x = reorder(rownames(pca2), -dim.1), y = dim.1)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + ggtitle("Dimension 1 loadings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# Dimension 2
ggplot(pca2, aes(x = reorder(rownames(pca2), -dim.2), y = dim.2)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + ggtitle("Dimension 2 loadings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
pcamix_data <- pca$scores[,1:6]

pcamix_data = cbind(hd_model1$target, pcamix_data)
colnames(pcamix_data) = c("target", "p1", "p2", "p3", "p4", "p5", "p6")

pcamix_data <- as.data.frame(pcamix_data)
pcamix_data <- pcamix_data %>% mutate(target = as.factor(target))

head(pcamix_data)
```


```{r}
pca_svm_model <- pcamix_data

set.seed(4)
train.samples_pca_svm <- pca_svm_model$target %>%
  createDataPartition(p = 0.7, list = FALSE)

train_data_pca_svm <- pca_svm_model[train.samples_pca_svm, ]
test_data_pca_svm <- pca_svm_model[-train.samples_pca_svm, ]
```


```{r}

pca_svm_lin <- train(target ~ ., data = train_data_pca_svm, method = "svmLinear", 
                 trControl = train_control_svm, tuneLength = 10)

pca_svm_lin
```

```{r}

pred_pca_svm <- predict(pca_svm_lin, train_data_pca_svm)

confusionMatrix(table(pred_pca_svm, train_data_pca_svm$target))
  
```

```{r}
grid_lin <- expand.grid(C = c(0.001, 0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2,5,10))

pca_svm_lin_grid<- train(target ~., data = train_data_pca_svm, method = "svmLinear", trControl = train_control_svm,
                     tuneGrid = grid_lin, tuneLength = 10)
pca_svm_lin_grid
plot(pca_svm_lin_grid)
```


```{r}
best_pca_svm_lin <- pca_svm_lin_grid
```


```{r}
pred_pca_svm_lin_best <- predict(best_pca_svm_lin, newdata = test_data_pca_svm)

confusionMatrix(pred_pca_svm_lin_best, test_data_pca_svm$target)
```

```{r}
tune_pca_svm_rad <- tune(svm, target ~ ., data = train_data_pca_svm, kernel = "radial", 
                         ranges = list(gamma = c(0, 0.0001, 0.001,0.01, 0.1, 0.3, 0.5, 0.75, 1), cost = c(5, 10, 50, 100)))

summary(tune_pca_svm_rad)
```

```{r}
best_pca_svm_rad <- tune_pca_svm_rad$best.model
```


```{r}
pred_pca_svm_rad_tune <- predict(best_pca_svm_rad, newdata = test_data_pca_svm)

confusionMatrix(pred_pca_svm_rad_tune, test_data_pca_svm$target)
```










