---
title: "Heart Disease EDA"
author: "Lucas Mendicino"
date: "11/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Packages
```{r}
library(tidyverse) 
library(DataExplorer)
library(gtsummary)
library(ggsci)
theme_gtsummary_journal(journal = "jama")

```

In this project, I will be using the Heart Disease dataset from the UC Irvine Machine Learning Repository. The dataset contains 14 features.

# Outline

1. Prepare the Data
2. Exploratory Data Analysis
3. Modeling
- Clustering (Principle Component Analysis, Hierarchical Clustering Analysis, K-Means)
- Logistic Regression
- Random Forest
- SVM
- 

# Prepare the Data

Let's load the data
```{r}
# Read datasets Cleveland_hd.csv into hd_data
hd <- read.csv("~/Desktop/UCI-Heart-Disease-Prediction/heart.csv")
```

Let's look at the structure and a few rows
```{r}
str(hd)
head(hd)
```
                                            
The variables include:

- age: age in years
- sex: (1 = male; 0 = female)
- cp: chest pain type (typical angina, atypical angina, non-anginal pain, or asymptomatic angina)
- trestbps: resting blood pressure (in mm Hg on admission to the hospital)
- chol: serum cholestoral in mg/dl
- fbs: Fasting blood sugar (< 120 mg/dl or > 120 mg/dl) (1 = true; 0 = false)
- restecg: resting electrocardiographic results (normal, ST-T wave abnormality, or left ventricular hypertrophy)
- thalach: Max. heart rate achieved during thalium stress test
- exang: Exercise induced angina (1 = yes; 0 = no)
- oldpeak: ST depression induced by exercise relative to rest
- slope: Slope of peak exercise ST segment (0 = upsloping, 1 = flat, or 2 = downsloping)
- ca: number of major vessels (0-3) colored by flourosopy 4 = NA
- thal: Thalium stress test result 3 = normal; 6 = fixed defect; 7 = reversable defect 0 = NA
- target: Heart disease status 1 or 0 (0 = heart disease 1 = asymptomatic)

It is important to note that for the target variable, 0 is assigned to heart disease and 1 is asymptomatic. Let's recode this.


```{r}
hd_new <- hd %>% mutate(target = recode(target, "0" = "1", "1" = "0"))
```


Let's tidy the dataset a bit more.

```{r}
hd_EDA <- hd_new %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
  ) %>%
# Recode the categorical variables as factors using the dplyr library.
  mutate(
    sex = case_when(
      sex == 0 ~ "female",
      sex == 1 ~ "male"
           ),
    fbs = case_when(
      fbs == 0 ~ "<=120",
      fbs == 1 ~ ">120"
            ),
    exang = case_when(
      exang == 0 ~ "no",
      exang == 1 ~ "yes"
            ),
    cp = case_when(
      cp == 3 ~ "typical angina",
      cp == 1 ~ "atypical angina",
      cp == 2 ~ "non-anginal pain",
      cp == 0 ~ "asymptomatic angina"
          ),
    restecg = case_when(
      restecg == 0 ~ "hypertrophy",
      restecg == 1 ~ "normal",
      restecg == 2 ~ "wave abnormality"
              ),
    target = case_when(
      target == 0 ~ "asymptomatic",
      target == 1 ~ "heart-disease"
              ),
    slope = case_when(
      slope == 2 ~ "upsloping",
      slope == 1 ~ "flat",
      slope == 0 ~ "downsloping"
    ),
    thal = case_when(
      thal == 1 ~ "fixed defect",
      thal == 2 ~ "normal",
      thal == 3 ~ "reversable defect"
    ),
    sex = as.factor(sex),
    fbs = as.factor(fbs),
    exang = as.factor(exang),
    cp = as.factor(cp),
    slope = as.factor(slope),
    ca = as.factor(ca),
    thal = as.factor(thal),
    restecg = as.factor(restecg),
    target = as.factor(target)
  )

```

```{r}
str(hd_EDA)
```

# Exploratory Data Analysis

Let's check out missing values
```{r}
plot_missing(hd_EDA)
```
The dataset is complete with no missing values.

Let's get summary statistics:

```{r, message = FALSE}
hd_EDA %>% tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
                    digits = all_continuous() ~ 2)
```







### Target: Whether patient has heart disease or not

```{r}
hd_EDA %>% group_by(target) %>% count()

ggplot(hd_EDA, aes(target, fill = target)) + geom_bar() + xlab("Disease Status") + ylab("Number of Patients")+ scale_color_jama()
```
### Age: age in years

```{r}
hd_EDA %>% summarise(Mean_Age = mean(age), Median_Age = median(age), SD_age = sd(age))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_Age = mean(age), Median_Age = median(age), SD_age = sd(age))
```

```{r}
ggplot(hd_EDA, aes(age, fill = target)) + geom_histogram(binwidth = 1) + labs(fill = "Disease Status", x = "Age", y = "Number of Patients") + facet_grid(~target) + scale_color_jama()
```

Does age have an effect? Let's use a t-test
```{r}
hd_age <- t.test(hd_EDA$age ~ hd_EDA$target)
print(hd_age)
```

```{r}
ggplot(data = hd_EDA, aes(x = target, y = age)) + geom_boxplot() + xlab("Heart Disease Diagnosis")
```

### Sex: Patient Sex

```{r}
hd_EDA %>% group_by(sex) %>% count()
```

```{r}
hd_EDA %>% group_by(target, sex) %>% count() 
```

Does sex have an effect? Let's use a chi-squared test
```{r}
hd_sex <- chisq.test(hd_EDA$sex, hd_EDA$target)
print(hd_sex)
```


```{r}
ggplot(data = hd_EDA, aes(x = sex, fill = target)) + geom_bar() + ylab("Number of Patients") + xlab("Heart Disease Diagnosis") + facet_grid(~target)+ scale_color_jama()
```

### cp: Chest Pain Type
- 0: asymptomatic
- 1: atypical angina
- 2: pain without relation to angina
- 3: typical angina


```{r}
hd_EDA %>% group_by(cp) %>% count()
```

```{r}
hd_EDA %>% group_by(target, cp) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(cp, fill = target)) + geom_bar() + labs(fill = "Disease Status", x = "Chest Pain Type", y = "Number of Patients") + 
  facet_grid(~target)+ scale_color_jama() + theme(axis.text.x = element_text(angle = 90))
```

```{r}
hd_cp <- chisq.test(hd_EDA$cp, hd_EDA$target)
print(hd_cp)
print(hd_cp$residuals)
```

### trestbps: Resting Blood Pressure in mmHg at admittance

```{r}
hd_EDA %>% summarise(Mean_bps = mean(trestbps), Median_bps = median(trestbps), SD_bps = sd(trestbps))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_bps = mean(trestbps), Median_bps = median(trestbps), SD_bps = sd(trestbps))
```

```{r}
ggplot(hd_EDA, aes(trestbps, fill = target)) + geom_histogram(binwidth = 3) + 
  labs(fill = "Disease Status", x = "Blood Pressure at Admittance (mmHg)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_bps <- t.test(hd_EDA$trestbps ~ hd_EDA$target)
print(hd_bps)
```

### chol: Cholesterol level in mg/dl

```{r}
hd_EDA %>% summarise(Mean_chol = mean(chol), Median_chol = median(chol), SD_chol = sd(chol))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_chol = mean(chol), Median_chol = median(chol), SD_chol = sd(chol))
```

```{r}
ggplot(hd_EDA, aes(chol, fill = target)) + geom_histogram(binwidth = 10) + 
  labs(fill = "Disease Status", x = "Cholesterol level (mg/dl)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_chol <- t.test(hd_EDA$chol ~ hd_EDA$target)
print(hd_chol)
```

### fbs: Whether Blood Sugar Level is greater than 120 mg/dl

```{r}
hd_EDA %>% group_by(fbs) %>% count()
```

```{r}
hd_EDA %>% group_by(target, fbs) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(fbs, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Blood Sugar Level (md/dl)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_fbs <- chisq.test(hd_EDA$fbs, hd_EDA$target)
print(hd_fbs)
```

### restecg: Resting Electrocardiogram 

- 0: probable left ventricular hypertrophy
- 1: normal
- 2: abnormalities in the T wave or ST segment


```{r}
hd_EDA %>% group_by(restecg) %>% count()
```

```{r}
hd_EDA %>% group_by(target, restecg) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(restecg, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Resting Electrocardiogram Result)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```


 use different test
```{r}
hd_ecg <- chisq.test(hd_EDA$restecg, hd_EDA$target)
print(hd_ecg)
print(hd_ecg$residuals)
```


### thalach: Maximum Heart Rate during Stress Test

```{r}
hd_EDA %>% summarise(Mean_thalach = mean(thalach), Median_thalach = median(thalach), SD_thalach = sd(thalach))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_thalach = mean(thalach), Median_thalach = median(thalach), SD_thalach = sd(thalach))
```

```{r}
ggplot(hd_EDA, aes(thalach, fill = target)) + geom_histogram(binwidth = 5) + 
  labs(fill = "Disease Status", x = "Maximu HR during Stress Test", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_thalach <- t.test(hd_EDA$thalach ~ hd_EDA$target)
print(hd_thalach)
```


```{r}
ggplot(data = hd_EDA, aes(x = target, y = thalach)) + geom_boxplot() + 
  xlab("Heart Disease Diagnosis") + ylab("Maximum Heart Rate") + 
  scale_color_jama()
```



### exang: Whether Patient had Angina During Exercise

```{r}
hd_EDA %>% group_by(exang) %>% count()
```

```{r}
hd_EDA %>% group_by(target, exang) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(exang, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Angina During Exercise)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```


```{r}
hd_exang <- chisq.test(hd_EDA$exang, hd_EDA$target)
print(hd_exang)
```

### oldpeak: Decrease of the ST segment during Exercise according to the same one on rest


```{r}
hd_EDA %>% summarise(Mean_oldpeak = mean(oldpeak), Median_oldpeak = median(oldpeak), SD_oldpeak = sd(oldpeak))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_oldpeak = mean(oldpeak), Median_oldpeak = median(oldpeak), SD_oldpeak = sd(oldpeak))
```

```{r}
ggplot(hd_EDA, aes(oldpeak, fill = target)) + geom_histogram(binwidth = 0.25) + 
  labs(fill = "Disease Status", x = "Decrease of ST Segment", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_oldpeak <- t.test(hd_EDA$oldpeak ~ hd_EDA$target)
print(hd_oldpeak)
```




### slope: Slope of the ST segment during the most demanding part of exercise

- 0: descending
- 1: flat
- 2: ascending


```{r}
hd_EDA %>% group_by(slope) %>% count()
```

```{r}
hd_EDA %>% group_by(target, slope) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(slope, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Slope of ST Segment)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```


```{r}
hd_exang <- chisq.test(hd_EDA$exang, hd_EDA$target)
print(hd_exang)
```

### thal: Results of the blood flow observed via radioactive dye

- 0: NULL (dropped from the dataset previously)
- 1: fixed defect (no blood flow in some part of the heart)
- 2: normal blood flow
- 3: reversible defect (a blood flow is observed but it is not normal)

```{r}
hd_EDA %>% group_by(thal) %>% count()
```

```{r}
hd_EDA %>% group_by(target, thal) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(thal, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Results of Blood Flow", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

need different test
```{r}
hd_exang <- chisq.test(hd_EDA$exang, hd_EDA$target)
print(hd_exang)
```


### ca: Number of main blood vessels coloured by radioactive dye


```{r}
hd_EDA %>% group_by(ca) %>% count()
```

```{r}
hd_EDA %>% group_by(target, ca) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(ca, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Number of Coloured Blood Vessels", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

need different test
```{r}
hd_ca <- chisq.test(hd_EDA$ca, hd_EDA$target)
print(hd_ca)
```


 Wilcoxon rank sum test; Pearson's Chi-squared test; Fisher's exact test 



```{r, message = FALSE}
hd_EDA %>% filter(thal != 0 & ca != 4) %>%
  tbl_summary(by = target) %>% add_p() 
```

age: Chi-Squared - heart disease older than asymptomatic
sex: Chi-Squared - More males with heart disease
cp: 



