---
title: "UCI Heart Disease Prediction"
author: "Lucas Mendicino"
date: "11/21/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages
```{r}
#EDA
library(tidyverse) 
library(DataExplorer)
library(gtsummary)
library(ggsci)
theme_gtsummary_journal(journal = "jama")
library(caret)
# Penalized Regression
library(glmnet)
# Random Forest
library(randomForest)
# Support Vector Machine
library(e1071)
# PCA
library(PCAmixdata)
library(factoextra)
library(kernlab)
library(mltools)
library(varhandle)
library(data.table)
```

In this project, I will be using the Heart Disease dataset from the UC Irvine Machine Learning Repository. The dataset contains 14 features.

# Outline

1. Prepare the Data
2. Exploratory Data Analysis
3. Modeling
- Clustering (Principle Component Analysis, Hierarchical Clustering Analysis, K-Means)
- Logistic Regression
- Random Forest
- SVM
- 

# Prepare the Data

Let's load the data
```{r}
# Read datasets Cleveland_hd.csv into hd_data
hd <- read.csv("~/Desktop/UCI-Heart-Disease-Prediction/heart.csv")
```

Let's look at the structure and a few rows
```{r}
str(hd)
head(hd)
```
                                            
The variables include:

- age: age in years
- sex: (1 = male; 0 = female)
- cp: chest pain type (typical angina, atypical angina, non-anginal pain, or asymptomatic angina)
- trestbps: resting blood pressure (in mm Hg on admission to the hospital)
- chol: serum cholestoral in mg/dl
- fbs: Fasting blood sugar (< 120 mg/dl or > 120 mg/dl) (1 = true; 0 = false)
- restecg: resting electrocardiographic results (normal, ST-T wave abnormality, or left ventricular hypertrophy)
- thalach: Max. heart rate achieved during thalium stress test
- exang: Exercise induced angina (1 = yes; 0 = no)
- oldpeak: ST depression induced by exercise relative to rest
- slope: Slope of peak exercise ST segment (0 = upsloping, 1 = flat, or 2 = downsloping)
- ca: number of major vessels (0-3) colored by flourosopy 4 = NA
- thal: Thalium stress test result 3 = normal; 6 = fixed defect; 7 = reversable defect 0 = NA
- target: Heart disease status 1 or 0 (0 = heart disease 1 = asymptomatic)

It is important to note that for the target variable, 0 is assigned to heart disease and 1 is asymptomatic. Let's recode this.

```{r}
hd_new <- hd %>% mutate(target = recode(target, "0" = "1", "1" = "0"))
```

Let's tidy the dataset a bit more.


```{r}
hd2 <- hd_new %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
  ) %>%
# Recode the categorical variables as factors using the dplyr library.
  mutate(
    sex = case_when(
      sex == 0 ~ "female",
      sex == 1 ~ "male"
           ),
    fbs = case_when(
      fbs == 0 ~ "<=120",
      fbs == 1 ~ ">120"
            ),
    exang = case_when(
      exang == 0 ~ "no",
      exang == 1 ~ "yes"
            ),
    cp = case_when(
      cp == 3 ~ "typical angina",
      cp == 1 ~ "atypical angina",
      cp == 2 ~ "non-anginal pain",
      cp == 0 ~ "asymptomatic angina"
          ),
    restecg = case_when(
      restecg == 0 ~ "hypertrophy",
      restecg == 1 ~ "normal",
      restecg == 2 ~ "wave abnormality"
              ),
    target = case_when(
      target == 0 ~ "asymptomatic",
      target == 1 ~ "heart-disease"
              ),
    slope = case_when(
      slope == 2 ~ "upsloping",
      slope == 1 ~ "flat",
      slope == 0 ~ "downsloping"
    ),
    thal = case_when(
      thal == 1 ~ "fixed defect",
      thal == 2 ~ "normal",
      thal == 3 ~ "reversable defect"
    ),
    sex = as.factor(sex),
    fbs = as.factor(fbs),
    exang = as.factor(exang),
    cp = as.factor(cp),
    slope = as.factor(slope),
    ca = as.factor(ca),
    thal = as.factor(thal),
    restecg = as.factor(restecg),
    target = as.factor(target)
  )

hd_EDA <- hd2
hd_model <- hd2

```


# Exploratory Data Analysis

Let's check out missing values
```{r}
plot_missing(hd_EDA)
```
The dataset is complete with no missing values.

Let's get summary statistics:

```{r, message = FALSE}
hd_EDA %>% tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
                    digits = all_continuous() ~ 2)
```


### Target: Whether patient has heart disease or not

```{r}
hd_EDA %>% group_by(target) %>% count()

ggplot(hd_EDA, aes(target, fill = target)) + geom_bar() + xlab("Disease Status") + ylab("Number of Patients")+ scale_color_jama()
```
### Age: age in years

```{r}
hd_EDA %>% summarise(Mean_Age = mean(age), Median_Age = median(age), SD_age = sd(age))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_Age = mean(age), Median_Age = median(age), SD_age = sd(age))
```

```{r}
ggplot(hd_EDA, aes(age, fill = target)) + geom_histogram(binwidth = 1) + labs(fill = "Disease Status", x = "Age", y = "Number of Patients") + facet_grid(~target) + scale_color_jama()
```

Does age have an effect? Let's use a t-test
```{r}
hd_age <- t.test(hd_EDA$age ~ hd_EDA$target)
print(hd_age)
```

```{r}
ggplot(data = hd_EDA, aes(x = target, y = age)) + geom_boxplot() + xlab("Heart Disease Diagnosis")
```

### Sex: Patient Sex

```{r}
hd_EDA %>% group_by(sex) %>% count()
```

```{r}
hd_EDA %>% group_by(target, sex) %>% count() 
```

Does sex have an effect? Let's use a chi-squared test
```{r}
hd_sex <- chisq.test(hd_EDA$sex, hd_EDA$target)
print(hd_sex)
```


```{r}
ggplot(data = hd_EDA, aes(x = sex, fill = target)) + geom_bar() + ylab("Number of Patients") + xlab("Heart Disease Diagnosis") + facet_grid(~target)+ scale_color_jama()
```

### cp: Chest Pain Type
- 0: asymptomatic
- 1: atypical angina
- 2: pain without relation to angina
- 3: typical angina


```{r}
hd_EDA %>% group_by(cp) %>% count()
```

```{r}
hd_EDA %>% group_by(target, cp) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(cp, fill = target)) + geom_bar() + labs(fill = "Disease Status", x = "Chest Pain Type", y = "Number of Patients") + 
  facet_grid(~target)+ scale_color_jama() + theme(axis.text.x = element_text(angle = 90))
```

```{r}
hd_cp <- chisq.test(hd_EDA$cp, hd_EDA$target)
print(hd_cp)
print(hd_cp$residuals)
```

### trestbps: Resting Blood Pressure in mmHg at admittance

```{r}
hd_EDA %>% summarise(Mean_bps = mean(trestbps), Median_bps = median(trestbps), SD_bps = sd(trestbps))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_bps = mean(trestbps), Median_bps = median(trestbps), SD_bps = sd(trestbps))
```

```{r}
ggplot(hd_EDA, aes(trestbps, fill = target)) + geom_histogram(binwidth = 3) + 
  labs(fill = "Disease Status", x = "Blood Pressure at Admittance (mmHg)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_bps <- t.test(hd_EDA$trestbps ~ hd_EDA$target)
print(hd_bps)
```

### chol: Cholesterol level in mg/dl

```{r}
hd_EDA %>% summarise(Mean_chol = mean(chol), Median_chol = median(chol), SD_chol = sd(chol))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_chol = mean(chol), Median_chol = median(chol), SD_chol = sd(chol))
```

```{r}
ggplot(hd_EDA, aes(chol, fill = target)) + geom_histogram(binwidth = 10) + 
  labs(fill = "Disease Status", x = "Cholesterol level (mg/dl)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_chol <- t.test(hd_EDA$chol ~ hd_EDA$target)
print(hd_chol)
```

### fbs: Whether Blood Sugar Level is greater than 120 mg/dl

```{r}
hd_EDA %>% group_by(fbs) %>% count()
```

```{r}
hd_EDA %>% group_by(target, fbs) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(fbs, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Blood Sugar Level (md/dl)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_fbs <- chisq.test(hd_EDA$fbs, hd_EDA$target)
print(hd_fbs)
```

### restecg: Resting Electrocardiogram 

- 0: probable left ventricular hypertrophy
- 1: normal
- 2: abnormalities in the T wave or ST segment


```{r}
hd_EDA %>% group_by(restecg) %>% count()
```

```{r}
hd_EDA %>% group_by(target, restecg) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(restecg, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Resting Electrocardiogram Result)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```


 use different test
```{r}
hd_ecg <- chisq.test(hd_EDA$restecg, hd_EDA$target)
print(hd_ecg)
print(hd_ecg$residuals)
```


### thalach: Maximum Heart Rate during Stress Test

```{r}
hd_EDA %>% summarise(Mean_thalach = mean(thalach), Median_thalach = median(thalach), SD_thalach = sd(thalach))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_thalach = mean(thalach), Median_thalach = median(thalach), SD_thalach = sd(thalach))
```

```{r}
ggplot(hd_EDA, aes(thalach, fill = target)) + geom_histogram(binwidth = 5) + 
  labs(fill = "Disease Status", x = "Maximu HR during Stress Test", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_thalach <- t.test(hd_EDA$thalach ~ hd_EDA$target)
print(hd_thalach)
```


```{r}
ggplot(data = hd_EDA, aes(x = target, y = thalach)) + geom_boxplot() + 
  xlab("Heart Disease Diagnosis") + ylab("Maximum Heart Rate") + 
  scale_color_jama()
```



### exang: Whether Patient had Angina During Exercise

```{r}
hd_EDA %>% group_by(exang) %>% count()
```

```{r}
hd_EDA %>% group_by(target, exang) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(exang, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Angina During Exercise)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```


```{r}
hd_exang <- chisq.test(hd_EDA$exang, hd_EDA$target)
print(hd_exang)
```

### oldpeak: Decrease of the ST segment during Exercise according to the same one on rest


```{r}
hd_EDA %>% summarise(Mean_oldpeak = mean(oldpeak), Median_oldpeak = median(oldpeak), SD_oldpeak = sd(oldpeak))
```

```{r}
hd_EDA %>% group_by(target) %>% summarise(Mean_oldpeak = mean(oldpeak), Median_oldpeak = median(oldpeak), SD_oldpeak = sd(oldpeak))
```

```{r}
ggplot(hd_EDA, aes(oldpeak, fill = target)) + geom_histogram(binwidth = 0.25) + 
  labs(fill = "Disease Status", x = "Decrease of ST Segment", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

```{r}
hd_oldpeak <- t.test(hd_EDA$oldpeak ~ hd_EDA$target)
print(hd_oldpeak)
```




### slope: Slope of the ST segment during the most demanding part of exercise

- 0: descending
- 1: flat
- 2: ascending


```{r}
hd_EDA %>% group_by(slope) %>% count()
```

```{r}
hd_EDA %>% group_by(target, slope) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(slope, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Slope of ST Segment)", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```


```{r}
hd_exang <- chisq.test(hd_EDA$exang, hd_EDA$target)
print(hd_exang)
```

### thal: Results of the blood flow observed via radioactive dye

- 0: NULL (dropped from the dataset previously)
- 1: fixed defect (no blood flow in some part of the heart)
- 2: normal blood flow
- 3: reversible defect (a blood flow is observed but it is not normal)

```{r}
hd_EDA %>% group_by(thal) %>% count()
```

```{r}
hd_EDA %>% group_by(target, thal) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(thal, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Results of Blood Flow", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

need different test
```{r}
hd_exang <- chisq.test(hd_EDA$exang, hd_EDA$target)
print(hd_exang)
```


### ca: Number of main blood vessels coloured by radioactive dye


```{r}
hd_EDA %>% group_by(ca) %>% count()
```

```{r}
hd_EDA %>% group_by(target, ca) %>% count() 
```


```{r}
ggplot(hd_EDA, aes(ca, fill = target)) + geom_bar() + 
  labs(fill = "Disease Status", x = "Number of Coloured Blood Vessels", y = "Number of Patients") +
  facet_grid(~target)+ scale_color_jama()
```

need different test
```{r}
hd_ca <- chisq.test(hd_EDA$ca, hd_EDA$target)
print(hd_ca)
```


 Wilcoxon rank sum test; Pearson's Chi-squared test; Fisher's exact test 



```{r, message = FALSE}
hd_EDA %>% filter(thal != 0 & ca != 4) %>%
  tbl_summary(by = target) %>% add_p() 
```

age: Chi-Squared - heart disease older than asymptomatic
sex: Chi-Squared - More males with heart disease
cp: 





# Machine Learning

## Training and Test sets



```{r}
hd_model <- na.omit(hd_model)

set.seed(44)
train.samples <- hd_model$target %>%
  createDataPartition(p = 0.7, list = FALSE)

train_data <- hd_model[train.samples, ]
test_data <- hd_model[-train.samples, ]
```

## Logistic Regression


full model
```{r}
log_reg <- glm(data = train_data, target ~ ., family = "binomial" )
summary(log_reg)
```



### LASSO

```{r}
x <- model.matrix(target ~ ., train_data)[,-1]
y <- ifelse(train_data$target == "heart-disease", 1 , 0)
```


```{r}
log_lasso <- glmnet(x, y, family = "binomial", alpha = 1, lambda = NULL)
plot(log_lasso)
```


Let's use misclassification error in 10-fold cross validation
```{r}
cv_lasso <- cv.glmnet(x, y, family = "binomial", alpha = 1, lambda = NULL, type.measure = "class")
plot(cv_lasso)
```

```{r}
cv_lasso$lambda.min
coef(cv_lasso, cv_lasso$lambda.min)
cv_lasso$lambda.1se
coef(cv_lasso, cv_lasso$lambda.1se)
```

```{r}
cv_lasso_auc <- cv.glmnet(x, y, family = "binomial", alpha = 1, lambda = NULL, type.measure = "auc", keep = TRUE)

rocs_l <- roc.glmnet(cv_lasso_auc$fit.preval, newy = y)
```

```{r}
best_l_min <- cv_lasso_auc$index["min",]

plot(rocs_l[[best_l_min]], type = "l")
invisible(sapply(rocs_l, lines, col="grey"))
lines(rocs_l[[best_l_min]], lwd = 2,col = "red")

```


```{r}
best_l_1se <- cv_lasso_auc$index["1se",]
plot(rocs_l[[best_l_1se]], type = "l")
invisible(sapply(rocs_l, lines, col="grey"))
lines(rocs_l[[best_l_1se]], lwd = 2,col = "blue")
```

LASSO Logistic Regression models
```{r}
lasso_model_min <- glmnet(x, y, alpha = 1, family = "binomial", 
                      lambda = cv_lasso$lambda.min)

lasso_model_1se <- glmnet(x, y, alpha = 1, family = "binomial", 
                      lambda = cv_lasso$lambda.1se)
```

Let's look at the performance of the two penalized models on the test data
```{r}
test_x <- model.matrix(target ~., test_data)[,-1]
test_y <- ifelse(test_data$target == "heart-disease", 1 , 0)


a_min <- assess.glmnet(lasso_model_min, newx = test_x, newy = test_y, s = "lambda.min")
a_1se <- assess.glmnet(lasso_model_1se, newx = test_x, newy = test_y, s = "lambda.1se")

do.call(rbind, Map(data.frame, Lasso_min = a_min, Lasso_1se = a_1se))
```



Let's closer at the LASSO model with the minimum lambda value
```{r}
roc_lasso_min_t <- roc.glmnet(lasso_model_min, newx = test_x, newy = test_y)
plot(roc_lasso_min_t, type = "l")
```


```{r}
min_pred <- predict(lasso_model_min, newx = test_x)

min_pred <- as.factor(ifelse(min_pred > 0.5, 1, 0))
cnf_min <- confusionMatrix(min_pred, as.factor(test_y), positive = "1")
cnf_min
```







Let's closer at the LASSO model with the 1 SE lambda value

```{r}
roc_lasso_1se_t <- roc.glmnet(lasso_model_1se, newx = test_x, newy = test_y)
plot(roc_lasso_1se_t, type = "l")
```


```{r}
onese_pred <- predict(lasso_model_1se, newx = test_x)

onese_pred <- as.factor(ifelse(onese_pred > 0.5, 1, 0))
cnf_1se <- confusionMatrix(onese_pred, as.factor(test_y), positive = "1")
cnf_1se
```



### Ridge

```{r}
log_ridge <- glmnet(x, y, family = "binomial", alpha = 0, lambda = NULL)
plot(log_ridge)
```


ridge model
```{r}
cv_ridge <- cv.glmnet(x, y, family = "binomial", alpha = 0, lambda = NULL, type.measure = "class")
```

```{r}
coef(cv_ridge, cv_ridge$lambda.min)
coef(cv_ridge, cv_ridge$lambda.1se)
```

```{r}
plot(cv_ridge)
```

```{r}
ridge_model_min <- glmnet(x, y, alpha = 0, family = "binomial", 
                      lambda = cv_ridge$lambda.min)

ridge_model_1se <- glmnet(x, y, alpha = 0, family = "binomial", 
                      lambda = cv_ridge$lambda.1se)
```



```{r}
test_x <- model.matrix(target ~., test_data)[,-1]
test_y <- ifelse(test_data$target == "heart-disease", 1 , 0)


ar_min <- assess.glmnet(ridge_model_min, newx = test_x, newy = test_y, s = "lambda.min")
ar_1se <- assess.glmnet(ridge_model_1se, newx = test_x, newy = test_y, s = "lambda.1se")

do.call(rbind, Map(data.frame, Ridge_min = ar_min, Ridge_1se = ar_1se))
```


```{r}
cv_ridge_auc_r <- cv.glmnet(x, y, family = "binomial", alpha = 0, lambda = NULL, type.measure = "auc", keep = TRUE)

rocs_r <- roc.glmnet(cv_ridge_auc_r$fit.preval, newy = y)
```

```{r}
best_r_min <- cv_ridge_auc_r$index["min",]
plot(rocs_r[[best_r_min]], type = "l")
invisible(sapply(rocs_r, lines, col="grey"))
lines(rocs_r[[best_r_min]], lwd = 2,col = "red")
```

```{r}
best_r_1se <- cv_ridge_auc_r$index["1se",]
plot(rocs_r[[best_r_1se]], type = "l")
invisible(sapply(rocs_r, lines, col="grey"))
lines(rocs_r[[best_r_1se]], lwd = 2,col = "blue")
```

```{r}
min_pred_r <- predict(ridge_model_min, newx = test_x)

min_pred_r <- as.factor(ifelse(min_pred_r > 0.5, 1, 0))
cnf_min_r <- confusionMatrix(min_pred_r, as.factor(test_y), positive = "1")
cnf_min_r
```





```{r}
onese_pred_r <- predict(ridge_model_1se, newx = test_x)

onese_pred_r <- as.factor(ifelse(onese_pred_r > 0.5, 1, 0))
cnf_1se_r <- confusionMatrix(onese_pred_r, as.factor(test_y), positive = "1")
cnf_1se_r
```

```{r}
do.call(rbind, Map(data.frame, Lasso_min = cnf_min$byClass, Lasso_1se = cnf_1se$byClass, Ridge_min = cnf_min_r$byClass, Ridge_1se = cnf_1se_r$byClass))

```




## Random Forest 



```{r}
train_x_rf <- train_data %>% select(-target)
train_y_rf <- train_data %>% select(target) %>% mutate(target = as.factor(target))

test_x_rf <- test_data %>% select(-target)
test_y_rf <- test_data %>% select(target) %>% mutate(target = as.factor(target))
```

```{r}
rf <- randomForest(target ~ ., data = train_data, importance = TRUE)
rf
```


```{r}
importance(rf)

varImpPlot(rf)

```


model tuning
```{r}
control_rf <- trainControl(method = "repeatedcv", number = 10, repeats = 3, search = "grid")

tune_grid <- expand.grid(mtry = (1:15))

rf_grid_search <- train(target ~ .,
                        data = train_data,
                        method = "rf",
                        metric = "Accuracy",
                        trControl = control_rf,
                        tuneGrid = tune_grid)

rf_grid_search
```

```{r}
rf_best <- rf_grid_search
plot(rf_best)

cnf_rf <- confusionMatrix(predict(rf_best, test_x_rf), test_y_rf[,1], positive = "heart-disease")
```

```{r}
ds_rf <- defaultSummary(data.frame(obs = test_y_rf[,1], 
                                   pred = predict(rf_best, test_x_rf)))

ds_rf
```


## Support Vector Machine



```{r}
hd_model_svm <- na.omit(hd_new)

missing_ca_indeces <- which(hd_model_svm$ca %in% 4)
missing_thal_indeces <-which(hd_model_svm$thal %in% 0)
restecg_indices <-which(hd_model_svm$restecg %in% 2)
missing_values_indeces <- c(missing_ca_indeces, missing_thal_indeces, restecg_indices)
hd_model_svm <- hd_model_svm[-missing_values_indeces, ]

hd_model_svm <- hd_model_svm %>% mutate(target = as.integer(target))
set.seed(44)
train.samples_svm <- hd_model_svm$target %>%
  createDataPartition(p = 0.7, list = FALSE)

train_data_svm <- hd_model_svm[train.samples_svm, ]
test_data_svm <- hd_model_svm[-train.samples_svm, ]
```


Let's scale the numeric variables and 1-hot encode the categorical ones

```{r}
train_data_svm <- train_data_svm %>% mutate(cp = as.factor(cp), restecg = as.factor(restecg),
                                            exang = as.factor(exang), slope = as.factor(slope),
                                            ca = as.factor(ca), thal = as.factor(thal))

test_data_svm <- test_data_svm %>% mutate(cp = as.factor(cp), restecg = as.factor(restecg),
                                            exang = as.factor(exang), slope = as.factor(slope),
                                            ca = as.factor(ca), thal = as.factor(thal))


train_data_svm[, c(1,4,5,8,10)] = scale(train_data_svm[, c(1,4,5,8,10)])
test_data_svm[, c(1,4,5,8,10)] = scale(test_data_svm[, c(1,4,5,8,10)])

dummy_train <- dummyVars(" ~ .", data = train_data_svm)
dummy_test <- dummyVars(" ~ .", data = test_data_svm)

train_data_svm <- data.frame(predict(dummy_train, newdata = train_data_svm))
test_data_svm <- data.frame(predict(dummy_test, newdata = test_data_svm))

train_data_svm$target <- as.factor(train_data_svm$target)
test_data_svm$target <- as.factor(test_data_svm$target)
```


```{r}
train_control_svm <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

svm_lin <- train(target ~ ., data = train_data_svm, method = "svmLinear", 
                 trControl = train_control_svm, tuneLength = 10)

svm_lin
```

```{r}
pred_svm <- predict(svm_lin, train_data_svm)
confusionMatrix(table(pred_svm, train_data_svm$target), positive = "1")
```

```{r}
grid_lin <- expand.grid(C = c(0.001, 0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2,5))

svm_lin_grid<- train(target ~., data = train_data_svm, method = "svmLinear", trControl = train_control_svm,
                     tuneGrid = grid_lin, tuneLength = 10)
svm_lin_grid
plot(svm_lin_grid)
```



```{r}
pred_svm_lin_grid <- predict(svm_lin_grid, newdata = test_data_svm)

cnf_svm_lin <- confusionMatrix(pred_svm_lin_grid, test_data_svm$target, positive = "1")
cnf_svm_lin
```



Radial
```{r}
svm_rad <- train(target ~ ., data = train_data_svm, method = "svmRadial", scale = FALSE,
                 trControl = train_control_svm, tuneLength = 10)

svm_rad
plot(svm_rad)
```

```{r}

pred_rad <- predict(svm_rad, train_data_svm)

confusionMatrix(table(pred_rad, train_data_svm$target))
  
```

```{r}
grid_rad <- expand.grid(sigma = c(0, 0.001, 0.01, 0.02, 0.1, 0.25, 0.5, 0.75,0.9),
 C = c(0.01, 0.05, 0.075, 0.1, 0.25, 0.5, 0.75, 1))

svm_rad_grid<- train(target ~., data = train_data_svm, method = "svmRadial", scale = FALSE, trControl = train_control_svm,
                     tuneGrid = grid_rad, tuneLength = 10)
svm_rad_grid
```


```{r}
plot(svm_rad_grid)
```


```{r}
pred_svm_rad_grid <- predict(svm_rad_grid, newdata = test_data_svm)

cnf_svm_rad <- confusionMatrix(pred_svm_rad_grid, test_data_svm$target, positive = "1")
cnf_svm_rad
```


## PCA then SVM

```{r}
hd_model1 <- hd_model %>% mutate(cp = as.factor(cp), restecg = as.factor(restecg),
                                            exang = as.factor(exang), slope = as.factor(slope),
                                            ca = as.factor(ca), thal = as.factor(thal), fbs = as.factor(fbs),
                                            sex = as.factor(sex)) 

hd_model2 <- hd_model1 %>% select(-c(target))
```

Using PCA of Mixed Data
```{r}
X.quanti <- splitmix(hd_model2)$X.quanti
X.quali <- splitmix(hd_model2)$X.quali

pca <- PCAmix(X.quanti, X.quali, ndim = 6, rename.level = TRUE)
summary(pca)
```

```{r}
pca$eig
```

```{r}
barplot(pca$eig[,2], ylim = c(0,20), las = 2)
```


```{r}
pca1 = data.frame(pca$ind$coord, hd_model1$target)
ggplot(pca1, aes(x = dim.1, y = dim.2, color = hd_model1.target)) + 
  geom_point() + 
  theme_bw() +
  guides(color = guide_legend(title = "Has heart \n disease?")) + 
  ggtitle("PCA of heart disease") +
  xlab(paste("Dimension 1", paste0("(", 
                                   round(pca$eig[1, 2], 2), 
                                   "%", ")"))) + 
  ylab(paste("Dimension 2", paste0("(", 
                                   round(pca$eig[2, 2], 2), 
                                   "%", ")")))
```

```{r}
pca2 = data.frame(pca$sqload)
# Dimension 1
ggplot(pca2, aes(x = reorder(rownames(pca2), -dim.1), y = dim.1)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + ggtitle("Dimension 1 loadings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# Dimension 2
ggplot(pca2, aes(x = reorder(rownames(pca2), -dim.2), y = dim.2)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + ggtitle("Dimension 2 loadings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
pcamix_data <- pca$scores[,1:6]

pcamix_data = cbind(hd_model1$target, pcamix_data)
colnames(pcamix_data) = c("target", "p1", "p2", "p3", "p4", "p5", "p6")

pcamix_data <- as.data.frame(pcamix_data)
pcamix_data <- pcamix_data %>% mutate(target = as.factor(target))

head(pcamix_data)
```


```{r}
pca_svm_model <- pcamix_data

set.seed(4)
train.samples_pca_svm <- pca_svm_model$target %>%
  createDataPartition(p = 0.7, list = FALSE)

train_data_pca_svm <- pca_svm_model[train.samples_pca_svm, ]
test_data_pca_svm <- pca_svm_model[-train.samples_pca_svm, ]
```


```{r}
pca_svm_lin <- train(target ~ ., data = train_data_pca_svm, method = "svmLinear", 
                 trControl = train_control_svm, tuneLength = 10)

pca_svm_lin
```

```{r}
pred_pca_svm <- predict(pca_svm_lin, train_data_pca_svm)

confusionMatrix(table(pred_pca_svm, train_data_pca_svm$target))
```

```{r}
grid_lin <- expand.grid(C = c(0.001, 0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2,3,4,5,10,20))

pca_svm_lin_grid<- train(target ~., data = train_data_pca_svm, method = "svmLinear", trControl = train_control_svm,
                     tuneGrid = grid_lin, tuneLength = 10)
pca_svm_lin_grid
plot(pca_svm_lin_grid)
```


```{r}
best_pca_svm_lin <- pca_svm_lin_grid
```


```{r}
pred_pca_svm_lin_best <- predict(best_pca_svm_lin, newdata = test_data_pca_svm)

cnf_pca_svm_lin <- confusionMatrix(pred_pca_svm_lin_best, test_data_pca_svm$target)
cnf_pca_svm_lin
```

```{r}
tune_pca_svm_rad <- tune(svm, target ~ ., data = train_data_pca_svm, kernel = "radial", 
                         ranges = list(gamma = c(0, 0.0001, 0.001,0.01, 0.1, 0.3, 0.5, 0.75, 1), cost = c(5, 10, 50, 100)))

summary(tune_pca_svm_rad)
```

```{r}
best_pca_svm_rad <- tune_pca_svm_rad$best.model
```


```{r}
pred_pca_svm_rad_tune <- predict(best_pca_svm_rad, newdata = test_data_pca_svm)

cnf_pca_svm_rad <- confusionMatrix(pred_pca_svm_rad_tune, test_data_pca_svm$target)
cnf_pca_svm_rad
```





## Compare Models







```{r}
do.call(rbind, Map(data.frame, Lasso_min = cnf_min$byClass, 
                   Lasso_1se = cnf_1se$byClass, 
                   Ridge_min = cnf_min_r$byClass, 
                   Ridge_1se = cnf_1se_r$byClass, 
                   RandomForest = cnf_rf$byClass, 
                   LinearSVM = cnf_svm_lin$byClass, 
                   RadialSVM = cnf_svm_rad$byClass,
                   PCASVMLinear = cnf_pca_svm_lin$byClass, 
                   PCASVMRadial = cnf_pca_svm_rad$byClass))

```



























