---
title: "Heart Disease Random Forest"
author: "Lucas Mendicino"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Packages
```{r}
library(tidyverse)
library(caret)
library(randomForest)

```

# Prepare the Data

Let's load the data
```{r}
# Read datasets Cleveland_hd.csv into hd_data
hd <- read.csv("~/Desktop/UCI-Heart-Disease-Prediction/heart.csv")
```

Let's look at the structure and a few rows
```{r}
str(hd)
head(hd)
```
                                            
The variables include:

- age: age in years
- sex: (1 = male; 0 = female)
- cp: chest pain type (typical angina, atypical angina, non-anginal pain, or asymptomatic angina)
- trestbps: resting blood pressure (in mm Hg on admission to the hospital)
- chol: serum cholestoral in mg/dl
- fbs: Fasting blood sugar (< 120 mg/dl or > 120 mg/dl) (1 = true; 0 = false)
- restecg: resting electrocardiographic results (normal, ST-T wave abnormality, or left ventricular hypertrophy)
- thalach: Max. heart rate achieved during thalium stress test
- exang: Exercise induced angina (1 = yes; 0 = no)
- oldpeak: ST depression induced by exercise relative to rest
- slope: Slope of peak exercise ST segment (0 = upsloping, 1 = flat, or 2 = downsloping)
- ca: number of major vessels (0-3) colored by flourosopy 4 = NA
- thal: Thalium stress test result 3 = normal; 6 = fixed defect; 7 = reversable defect 0 = NA
- target: Heart disease status 1 or 0 (0 = heart disease 1 = asymptomatic)

It is important to note that for the target variable, 0 is assigned to heart disease and 1 is asymptomatic. Let's recode this.


```{r}
hd_new <- hd %>% mutate(target = recode(target, "0" = "1", "1" = "0"))
```


Let's tidy the dataset a bit more.

```{r}
hd_model <- hd_new %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
  ) %>%
# Recode the categorical variables as factors using the dplyr library.
  mutate(
    sex = case_when(
      sex == 0 ~ "female",
      sex == 1 ~ "male"
           ),
    fbs = case_when(
      fbs == 0 ~ "<=120",
      fbs == 1 ~ ">120"
            ),
    exang = case_when(
      exang == 0 ~ "no",
      exang == 1 ~ "yes"
            ),
    cp = case_when(
      cp == 3 ~ "typical angina",
      cp == 1 ~ "atypical angina",
      cp == 2 ~ "non-anginal pain",
      cp == 0 ~ "asymptomatic angina"
          ),
    restecg = case_when(
      restecg == 0 ~ "hypertrophy",
      restecg == 1 ~ "normal",
      restecg == 2 ~ "wave abnormality"
              ),
    target = case_when(
      target == 0 ~ "asymptomatic",
      target == 1 ~ "heart-disease"
              ),
    slope = case_when(
      slope == 2 ~ "upsloping",
      slope == 1 ~ "flat",
      slope == 0 ~ "downsloping"
    ),
    thal = case_when(
      thal == 1 ~ "fixed defect",
      thal == 2 ~ "normal",
      thal == 3 ~ "reversable defect"
    ),
    sex = as.factor(sex),
    fbs = as.factor(fbs),
    exang = as.factor(exang),
    cp = as.factor(cp),
    slope = as.factor(slope),
    ca = as.factor(ca),
    thal = as.factor(thal),
    restecg = as.factor(restecg),
    target = as.factor(target)
  )

```



## Training and Test sets


```{r}
hd_model <- na.omit(hd_model)

set.seed(44)
train.samples <- hd_model$target %>%
  createDataPartition(p = 0.7, list = FALSE)

train_data <- hd_model[train.samples, ]
test_data <- hd_model[-train.samples, ]

train_x <- train_data %>% select(-target)
train_y <- train_data %>% select(target) %>% mutate(target = as.factor(target))

test_x <- test_data %>% select(-target)
test_y <- test_data %>% select(target) %>% mutate(target = as.factor(target))

```




```{r}
rf <- randomForest(target ~ ., data = train_data, importance = TRUE)
rf
```


```{r}
importance(rf)

varImpPlot(rf)

```


model tuning
```{r}
control_rf <- trainControl(method = "repeatedcv", number = 10, repeats = 3, search = "grid")

tune_grid <- expand.grid(mtry = (1:15))

rf_grid_search <- train(target ~ .,
                        data = train_data,
                        method = "rf",
                        metric = "Accuracy",
                        trControl = control_rf,
                        tuneGrid = tune_grid)

rf_grid_search
```

```{r}
rf_best <- rf_grid_search
plot(rf_best)

confusionMatrix(predict(rf_best, test_x), test_y[,1], positive = "heart-disease")
```

```{r}
ds_rf <- defaultSummary(data.frame(obs = test_y[,1], 
                                   pred = predict(rf_best, test_x)))

ds_rf
```

