---
title: "Heart Disease Support Vector Machine"
author: "Lucas Mendicino"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Support Vector Machine


# Load Packages
```{r}
library(tidyverse)
library(caret)
library(e1071)
```

# Prepare the Data

Let's load the data
```{r}
# Read datasets Cleveland_hd.csv into hd_data
hd <- read.csv("~/Desktop/UCI-Heart-Disease-Prediction/heart.csv")
```

Let's look at the structure and a few rows
```{r}
str(hd)
head(hd)
```
                                            
The variables include:

- age: age in years
- sex: (1 = male; 0 = female)
- cp: chest pain type (typical angina, atypical angina, non-anginal pain, or asymptomatic angina)
- trestbps: resting blood pressure (in mm Hg on admission to the hospital)
- chol: serum cholestoral in mg/dl
- fbs: Fasting blood sugar (< 120 mg/dl or > 120 mg/dl) (1 = true; 0 = false)
- restecg: resting electrocardiographic results (normal, ST-T wave abnormality, or left ventricular hypertrophy)
- thalach: Max. heart rate achieved during thalium stress test
- exang: Exercise induced angina (1 = yes; 0 = no)
- oldpeak: ST depression induced by exercise relative to rest
- slope: Slope of peak exercise ST segment (0 = upsloping, 1 = flat, or 2 = downsloping)
- ca: number of major vessels (0-3) colored by flourosopy 4 = NA
- thal: Thalium stress test result 3 = normal; 6 = fixed defect; 7 = reversable defect 0 = NA
- target: Heart disease status 1 or 0 (0 = heart disease 1 = asymptomatic)

It is important to note that for the target variable, 0 is assigned to heart disease and 1 is asymptomatic. Let's recode this.


```{r}
hd_new <- hd %>% mutate(target = recode(target, "0" = "1", "1" = "0"))
```


Let's tidy the dataset a bit more.

```{r}
hd_model <- hd_new %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
  ) %>%
# Recode the categorical variables as factors using the dplyr library.
  mutate(
    sex = case_when(
      sex == 0 ~ "female",
      sex == 1 ~ "male"
           ),
    fbs = case_when(
      fbs == 0 ~ "<=120",
      fbs == 1 ~ ">120"
            ),
    exang = case_when(
      exang == 0 ~ "no",
      exang == 1 ~ "yes"
            ),
    cp = case_when(
      cp == 3 ~ "typical angina",
      cp == 1 ~ "atypical angina",
      cp == 2 ~ "non-anginal pain",
      cp == 0 ~ "asymptomatic angina"
          ),
    restecg = case_when(
      restecg == 0 ~ "hypertrophy",
      restecg == 1 ~ "normal",
      restecg == 2 ~ "wave abnormality"
              ),
    target = case_when(
      target == 0 ~ "asymptomatic",
      target == 1 ~ "heart-disease"
              ),
    slope = case_when(
      slope == 2 ~ "upsloping",
      slope == 1 ~ "flat",
      slope == 0 ~ "downsloping"
    ),
    thal = case_when(
      thal == 1 ~ "fixed defect",
      thal == 2 ~ "normal",
      thal == 3 ~ "reversable defect"
    ),
    sex = as.factor(sex),
    fbs = as.factor(fbs),
    exang = as.factor(exang),
    cp = as.factor(cp),
    slope = as.factor(slope),
    ca = as.factor(ca),
    thal = as.factor(thal),
    restecg = as.factor(restecg),
    target = as.factor(target)
  )

```



## Training and Test sets



```{r}
hd_model_svm <- na.omit(hd_new)

missing_ca_indeces <- which(hd_model_svm$ca %in% 4)
missing_thal_indeces <-which(hd_model_svm$thal %in% 0)
restecg_indices <-which(hd_model_svm$restecg %in% 2)
missing_values_indeces <- c(missing_ca_indeces, missing_thal_indeces, restecg_indices)
hd_model_svm <- hd_model_svm[-missing_values_indeces, ]

hd_model_svm <- hd_model_svm %>% mutate(target = as.integer(target))
set.seed(44)
train.samples <- hd_model_svm$target %>%
  createDataPartition(p = 0.7, list = FALSE)

train_data <- hd_model_svm[train.samples, ]
test_data <- hd_model_svm[-train.samples, ]
```



Let's scale the numeric variables and 1-hot encode the categorical ones


```{r}
train_data_svm <- train_data
test_data_svm <- test_data

train_data_svm <- train_data_svm %>% mutate(cp = as.factor(cp), restecg = as.factor(restecg),
                                            exang = as.factor(exang), slope = as.factor(slope),
                                            ca = as.factor(ca), thal = as.factor(thal))

test_data_svm <- test_data_svm %>% mutate(cp = as.factor(cp), restecg = as.factor(restecg),
                                            exang = as.factor(exang), slope = as.factor(slope),
                                            ca = as.factor(ca), thal = as.factor(thal))


train_data_svm[, c(1,4,5,8,10)] = scale(train_data_svm[, c(1,4,5,8,10)])
test_data_svm[, c(1,4,5,8,10)] = scale(test_data_svm[, c(1,4,5,8,10)])

dummy_train <- dummyVars(" ~ .", data = train_data_svm)
dummy_test <- dummyVars(" ~ .", data = test_data_svm)

train_data_svm <- data.frame(predict(dummy_train, newdata = train_data_svm))
test_data_svm <- data.frame(predict(dummy_test, newdata = test_data_svm))

train_data_svm$target <- as.factor(train_data_svm$target)
test_data_svm$target <- as.factor(test_data_svm$target)
```






```{r}
train_control_svm <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

svm_lin <- train(target ~ ., data = train_data_svm, method = "svmLinear", 
                 trControl = train_control_svm, tuneLength = 10)

svm_lin
```

```{r}

pred_svm <- predict(svm_lin, train_data_svm)

confusionMatrix(table(pred_svm, train_data_svm$target), positive = "1")
  
```

```{r}
grid_lin <- expand.grid(C = c(0.001, 0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2,5))

svm_lin_grid<- train(target ~., data = train_data_svm, method = "svmLinear", trControl = train_control_svm,
                     tuneGrid = grid_lin, tuneLength = 10)
svm_lin_grid
plot(svm_lin_grid)
```



```{r}
pred_svm_lin_grid <- predict(svm_lin_grid, newdata = test_data_svm)

confusionMatrix(pred_svm_lin_grid, test_data_svm$target, positive = "1")
```



Radial
```{r}
svm_rad <- train(target ~ ., data = train_data_svm, method = "svmRadial", scale = FALSE,
                 trControl = train_control_svm, tuneLength = 10)

svm_rad
plot(svm_rad)
```

```{r}

pred_rad <- predict(svm_rad, train_data_svm)

confusionMatrix(table(pred_rad, train_data_svm$target))
  
```

```{r}
grid_rad <- expand.grid(sigma = c(0, 0.001, 0.01, 0.02, 0.1, 0.25, 0.5, 0.75,0.9),
 C = c(0.01, 0.05, 0.075, 0.1, 0.25, 0.5, 0.75, 1))

svm_rad_grid<- train(target ~., data = train_data_svm, method = "svmRadial", scale = FALSE, trControl = train_control_svm,
                     tuneGrid = grid_rad, tuneLength = 10)
svm_rad_grid
```


```{r}
plot(svm_rad_grid)
```


```{r}
pred_svm_rad_grid <- predict(svm_rad_grid, newdata = test_data_svm)

confusionMatrix(pred_svm_rad_grid, test_data_svm$target, positive = "1")
```
