---
title: "Heart Disease Logistic Regression"
author: "Lucas Mendicino"
date: "11/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Packages
```{r}
library(tidyverse) 
library(caret)
library(glmnet)
```

```{r}
# Read datasets Cleveland_hd.csv into hd_data
hd <- read.csv("~/Desktop/UCI-Heart-Disease-Prediction/heart.csv")
```

Let's look at the structure and a few rows
```{r}
str(hd)
head(hd)
```
                                            
The variables include:

- age: age in years
- sex: (1 = male; 0 = female)
- cp: chest pain type (typical angina, atypical angina, non-anginal pain, or asymptomatic angina)
- trestbps: resting blood pressure (in mm Hg on admission to the hospital)
- chol: serum cholestoral in mg/dl
- fbs: Fasting blood sugar (< 120 mg/dl or > 120 mg/dl) (1 = true; 0 = false)
- restecg: resting electrocardiographic results (normal, ST-T wave abnormality, or left ventricular hypertrophy)
- thalach: Max. heart rate achieved during thalium stress test
- exang: Exercise induced angina (1 = yes; 0 = no)
- oldpeak: ST depression induced by exercise relative to rest
- slope: Slope of peak exercise ST segment (0 = upsloping, 1 = flat, or 2 = downsloping)
- ca: number of major vessels (0-3) colored by flourosopy 4 = NA
- thal: Thalium stress test result 3 = normal; 6 = fixed defect; 7 = reversable defect 0 = NA
- target: Heart disease status 1 or 0 (0 = heart disease 1 = asymptomatic)

It is important to note that for the target variable, 0 is assigned to heart disease and 1 is asymptomatic. Let's recode this.


```{r}
hd_new <- hd %>% mutate(target = recode(target, "0" = "1", "1" = "0"))
```


Let's tidy the dataset a bit more.

```{r}
hd_model <- hd_new %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
  ) %>%
# Recode the categorical variables as factors using the dplyr library.
  mutate(
    sex = case_when(
      sex == 0 ~ "female",
      sex == 1 ~ "male"
           ),
    fbs = case_when(
      fbs == 0 ~ "<=120",
      fbs == 1 ~ ">120"
            ),
    exang = case_when(
      exang == 0 ~ "no",
      exang == 1 ~ "yes"
            ),
    cp = case_when(
      cp == 3 ~ "typical angina",
      cp == 1 ~ "atypical angina",
      cp == 2 ~ "non-anginal pain",
      cp == 0 ~ "asymptomatic angina"
          ),
    restecg = case_when(
      restecg == 0 ~ "hypertrophy",
      restecg == 1 ~ "normal",
      restecg == 2 ~ "wave abnormality"
              ),
    target = case_when(
      target == 0 ~ "asymptomatic",
      target == 1 ~ "heart-disease"
              ),
    slope = case_when(
      slope == 2 ~ "upsloping",
      slope == 1 ~ "flat",
      slope == 0 ~ "downsloping"
    ),
    thal = case_when(
      thal == 1 ~ "fixed defect",
      thal == 2 ~ "normal",
      thal == 3 ~ "reversable defect"
    ),
    sex = as.factor(sex),
    fbs = as.factor(fbs),
    exang = as.factor(exang),
    cp = as.factor(cp),
    slope = as.factor(slope),
    ca = as.factor(ca),
    thal = as.factor(thal),
    restecg = as.factor(restecg),
    target = as.factor(target)
  )

```


## Training and Test sets


```{r}
hd_model <- na.omit(hd_model)

set.seed(44)
train.samples <- hd_model$target %>%
  createDataPartition(p = 0.7, list = FALSE)

train_data <- hd_model[train.samples, ]
test_data <- hd_model[-train.samples, ]
```

## Logistic Regression


full model
```{r}
log_reg <- glm(data = train_data, target ~ ., family = "binomial" )
summary(log_reg)
```



### LASSO

```{r}
x <- model.matrix(target ~ ., train_data)[,-1]
y <- ifelse(train_data$target == "heart-disease", 1 , 0)
```


```{r}
log_lasso <- glmnet(x, y, family = "binomial", alpha = 1, lambda = NULL)
plot(log_lasso)
```


Let's use misclassification error in 10-fold cross validation
```{r}
cv_lasso <- cv.glmnet(x, y, family = "binomial", alpha = 1, lambda = NULL, type.measure = "class")
plot(cv_lasso)
```

```{r}
cv_lasso$lambda.min
coef(cv_lasso, cv_lasso$lambda.min)
cv_lasso$lambda.1se
coef(cv_lasso, cv_lasso$lambda.1se)
```

```{r}
cv_lasso_auc <- cv.glmnet(x, y, family = "binomial", alpha = 1, lambda = NULL, type.measure = "auc", keep = TRUE)

rocs_l <- roc.glmnet(cv_lasso_auc$fit.preval, newy = y)
```

```{r}
best_l_min <- cv_lasso_auc$index["min",]

plot(rocs_l[[best_l_min]], type = "l")
invisible(sapply(rocs_l, lines, col="grey"))
lines(rocs_l[[best_l_min]], lwd = 2,col = "red")

```


```{r}
best_l_1se <- cv_lasso_auc$index["1se",]
plot(rocs_l[[best_l_1se]], type = "l")
invisible(sapply(rocs_l, lines, col="grey"))
lines(rocs_l[[best_l_1se]], lwd = 2,col = "blue")
```



```{r}
lasso_model_min <- glmnet(x, y, alpha = 1, family = "binomial", 
                      lambda = cv_lasso$lambda.min)

lasso_model_1se <- glmnet(x, y, alpha = 1, family = "binomial", 
                      lambda = cv_lasso$lambda.1se)
```


```{r}
test_x <- model.matrix(target ~., test_data)[,-1]
test_y <- ifelse(test_data$target == "heart-disease", 1 , 0)


a_min <- assess.glmnet(lasso_model_min, newx = test_x, newy = test_y, s = "lambda.min")
a_1se <- assess.glmnet(lasso_model_1se, newx = test_x, newy = test_y, s = "lambda.1se")

do.call(rbind, Map(data.frame, Lasso_min = a_min, Lasso_1se = a_1se))
```

Let's look at the AUC
```{r}
roc_lasso_min_t <- roc.glmnet(lasso_model_min, newx = test_x, newy = test_y)
plot(roc_lasso_min_t, type = "l")
```

```{r}
cnf_min <- confusion.glmnet(lasso_model_min, newx = test_x, newy = test_y)
print(cnf_min)
```

```{r}
# Make prediction on test data
test_x <- model.matrix(target ~., test_data)[,-1]
probabilities_min <- lasso_model_min %>% predict(newx = test_x)
predicted.classes_min <- ifelse(probabilities_min > 0.5, "heart-disease", "asymptomatic")

# Model accuracy rate
observed.classes <- test_data$target
mean(predicted.classes_min == observed.classes)
```

```{r}
roc_lasso_1se_t <- roc.glmnet(lasso_model_1se, newx = test_x, newy = test_y)
plot(roc_lasso_1se_t, type = "l")
```

```{r}
cnf_1se <- confusion.glmnet(lasso_model_1se, newx = test_x, newy = test_y)
print(cnf_1se)
```


```{r}
# Make prediction on test data
test_x <- model.matrix(target ~., test_data)[,-1]
probabilities_1se <- lasso_model_1se %>% predict(newx = test_x)
predicted.classes_1se <- ifelse(probabilities_1se > 0.5, "heart-disease", "asymptomatic")

# Model accuracy rate
observed.classes <- test_data$target
mean(predicted.classes_1se == observed.classes)
```

### Ridge

```{r}
log_ridge <- glmnet(x, y, family = "binomial", alpha = 0, lambda = NULL)
plot(log_ridge)
```


ridge model
```{r}
cv_ridge <- cv.glmnet(x, y, family = "binomial", alpha = 0, lambda = NULL, type.measure = "class")
```

```{r}
coef(cv_ridge, cv_ridge$lambda.min)
coef(cv_ridge, cv_ridge$lambda.1se)
```

```{r}
plot(cv_ridge)
```

```{r}
ridge_model <- glmnet(x, y, alpha = 0, family = "binomial", 
                      lambda = cv_ridge$lambda.min)
```



```{r}
test_x <- model.matrix(target ~., test_data)[,-1]
test_y <- ifelse(test_data$target == "heart-disease", 1 , 0)


assess.glmnet(ridge_model, newx = test_x, newy = test_y, s = "lambda.min")


```

```{r}
cv_ridge_auc <- cv.glmnet(x, y, family = "binomial", alpha = 0, lambda = NULL, type.measure = "auc", keep = TRUE)

rocs_r <- roc.glmnet(cv_ridge_auc$fit.preval, newy = y)
```

```{r}
best_r <- cv_ridge_auc$index["min",]
plot(rocs_r[[best_r]], type = "l")
invisible(sapply(rocs_r, lines, col="grey"))
lines(rocs_r[[best_r]], lwd = 2,col = "red")
```

```{r}
cnf_r <- confusion.glmnet(ridge_model, newx = test_x, newy = test_y)
print(cnf_r)
```



```{r}

# Make prediction on test data
test_x <- model.matrix(target ~., test_data)[,-1]
probabilities_r <- ridge_model %>% predict(newx = test_x)
predicted.classes_r <- ifelse(probabilities_r > 0.5, "heart-disease", "asymptomatic")

# Model accuracy rate
observed.classes_r <- test_data$target
mean(predicted.classes_r == observed.classes_r)

```






